---
title: "DDP Final Project Pitch Presentation"
author: "Pratik A. Khare"
date: "`r Sys.Date()`"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

- This app predicts the probability of India winning a cricket match using 
  **multinomial logistic regression** trained on historical match results

- The model takes three user-selected inputs namely, **opponent, venue, and format** 
  (Test / ODI / T20), and returns the predicted probability of a Win, Loss, or Draw

- The model is simplistic in nature, considering only a few key variables. 
  Additionally, it is trained on historical data. Past performance is not 
  a guarantee of future results

## ML algorithm used and evaluation

The app uses a **multinomial logistic regression** model via the `multinom()` 
function from the `nnet` package. The model is trained on historical India 
match results, predicting outcome (Win/Loss/Draw) from opponent, venue, and format.
```{r model, echo = TRUE}
library(readxl)
library(nnet)

data <- read_excel("data/Ind vs opponent results.xlsx")
df <- subset(data, Result != "NR")
df$Result[df$Result %in% c("Tied","Draw")] <- "Draw"
df$Result <- factor(df$Result)
df$Opponent <- factor(df$Opponent)
df$Venue <- factor(df$Venue)
df$Format <- factor(df$Format)

model <- multinom(Result ~ Opponent + Venue + Format, data = df)
summary(model)
```

## Results in graphical format

```{r, echo = FALSE}
library(ggplot2)
predicted <- predict(model, df)
cm <- as.data.frame(table(Predicted = predicted, Actual = df$Result))

ggplot(cm, aes(x = Actual, y = Predicted, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq), size = 6) +
  scale_fill_gradient(low = "white", high = "steelblue") +
  labs(title = "Confusion Matrix — India Match Outcome Prediction")
```

Overall accuracy is moderate; the model performs best on Win predictions but struggles with Draw classification due to limited predictor variables. 
  
## Note on Data Collection

- Match results data was manually compiled from **ESPN Cricinfo**, covering 
  India's international matches across Test, ODI, and T20 formats against 
  five major opponents: Australia, England, South Africa, Pakistan, and West Indies

- Each record contains four variables: **Opponent, Venue, Format, and Result**. 
  Venue is categorized as Home, Away, or Neutral. No-result (NR) matches 
  were excluded, and Tied/Drawn matches were merged into a single **Draw** category

- The dataset is limited in scope; only five opponents are included, and 
  historical factors such as team composition, player form, and pitch 
  conditions are not captured. Users should interpret predictions accordingly